The Types and Driving Forces of Modifications in the Evolution of an API

### 2015-05-12 thesis/paper.tex
More changes to results

\subsection{API Design} \label{api_design}
API design is notoriously difficult, as a myriad of design and performance decisions must be taken into consideration when creating APIs \cite{afonso2012evaluating, bloch2008effective, stylos2006comparing}.

 Examples of such design decisions include how to structure an object's constructor parameters or if the API should display errors at compilation or at runtime \cite{stylos2006comparing}. More trivial design problems, such as assigning names to API features or correctly naming user-defined types, can have a significant impact on the usability of an API \cite{shi2011empirical}. When facing such design challenges, the following four factors are important to consider: a) The API must be understandable through good documentation, b) the API must not be overly abstract, c) the API must be reusable and d) the API must be easy to learn \cite{shi2011empirical}. One of the most important qualities in an API is that the intent of the API must be clear to the user \cite{stylos2006comparing, shi2011empirical}. The design decisions reached during development of an API will affect the overall usability of the API. Measuring such an effect can be done by investigating the twelve cognitive dimensions that are impacted by interactions between the API and its users \cite{clarke2004measuring}.

@@

%%Lehman's laws (!!!improve later)
Software evolution theory suggests that changes to software must obey one of eight laws \cite{lehman1980programs}. The laws, known as Lehman's laws, are: 1) Continuing Change, 2) Increasing Complexity, 3) Self Regulation, 4) Conservation of Organisational Stability, 5) Conservation of Familiarity, 6) Continuing Growth, 7) Declining Quality and 8) Feedback System. The laws are said to only apply to E-type systems, which are programs that solve problems or address applications in the real world.

@@

\subsection{API Evolution} \label{related_work}
The analysis of APIs in the context of software evolution has primarily focused on APIs that are part of large programming languages, such as Java \cite{hou2011exploring, shi2011empirical} and Smalltalk \cite{robbes2012developers}. Attempts to uncover the intents behind the changes made in the AWT and Swing Java libraries have also been undertaken \cite{hou2011exploring}, leading to suggestions that the use of a strong architecture is vital for ensuring the successful evolution of an API \cite{hou2011exploring}.

@@

%%Possibly add a paragraph here to motivate the gap more

@@

@@

%Reasons for choosing case study
The reason the selected methodology was used is that it is essential to study the phenomenon of API change in its natural context. The applicability of case studies in such scenarios is supported by existing literature \cite{benbasat1987case, runeson2009guidelines, yin2013case, robson2002real}. An alternative approach that was considered was design research, but if a prototype API was to be used instead of one which is tried and tested in an industry setting, the study would lack real-life context \cite{runeson2009guidelines}. An additional motivation for why the case study approach was used is that there is little existing research conducted in the area of motives behind API change and that input from the industry is vital to the success of the research, especially to be able to answer \textbf{RQ2}.

@@

The process used for examining the code and documentation was inspired by grounded theory analysis \cite{seaman1999qualitative}, as it is recommended for hypothesis generating studies \cite{runeson2009guidelines, seaman1999qualitative}.

@@

\smallskip \noindent
\textbf{Construct validity  } Threats to construct validity have largely been mitigated by structuring the results of the study around already established and accepted software evolution theories \cite{chapin2001types, lehman1980programs}. Due to restricted access to the company's source code, we were not able to include versions prior to version 1.4 in our analysis. This means that the evolution of the API was studied during a rather short period of its lifetime. It was also not possible to analyse the implementation of the API, something which might have contributed to more accurate predictions of the motives behind the changes. We consider these factors to be the remaining threats to construct validity.

@@

\smallskip \noindent
\textbf{Reliability  } Potential threats to the reliability of our study have, to a great extent, been reduced by conducting the study according to an accredited guide to case studies \cite{runeson2009guidelines}. In addition to that, strategies from additional well-established papers \cite{andersson2007spiral, seaman1999qualitative, robson2002real} have been used to increase the trustworthiness of the data gathering and data analysis. The analysis was based around well-defined theories that are cemented in academia, to further increase the reproducibility of the study. The semi-structured nature of the interview, might threaten the reproducibility; however, large parts of it were structured around existing theories.

@@

\subsubsection{Categories of Maintenance}

The majority of changes between version 1.4 and version 2.0 have been additions to functionality of the API. Such additions include added functionality to mechanically and digitally control the camera, to utilise additional storage devices, to allow serial connectivity and to allow audio analysis on the cameras.

Many of the changes found many deprecations which were later re-implemented as new modules. We have grouped these kinds of changes into a single type instead of two. Such examples include the deprecation and re-implementation of an event system, dynamic web page generation and configuration utilities.

The remainder of the changes were defined as general trends. These changes were not individual, but were instead larger changes that effected the whole API. Such examples include a shift towards using interfaces to access features, improved maintainability, a different error-handling approach and the inclusion of additional 3\textit{rd} party libraries.

@@

\smallskip \noindent
\textbf{Corrective  }
%Define what kinds of changes these are
Corrective changes are identified as types of change that fix broken functionality.
%Data and explanation
No corrective changes were identified between V1.0 and V2.0. There are two main reasons that we did not find any changes of this type. Firstly, the majority of changes were additions or re-implementations of functionality. This meant that most of the existing code was either removed, replaced or deprecated resulting in little to no obvious fixes. Secondly, we only had access to the public facing API and not the implementation itself. We therefore could not inspect the inner workings of the functionality that has remained between versions to determine if any corrective change took place.
%Conclusion
We therefore conclude that the evolution of the API was not driven by corrective changes.
\smallskip \noindent
\textbf{Reductive  }
%Definition
Reductive changes are defined as changes that remove or reduce functionality. This is distinctive from a deprecation as reductive types of change remove functionality completely.
%Data & Explanation
Only a single reductive type of change was identified. This reductive change removed functionality that provided the ability to buffer individual images for closer analysis by a user. Given that it is more common to refactor APIs and not remove functionality \cite{dig2005role} \cite{xing2006refactoring}, this was a surprising result. 
%Don't know if I should discuss the finding in this much depth here?
%If OK - Talk more about the business side of things, and how we couldn't identify this change ourselves
It was revealed during the interview that the functionality was removed because the case company no longer wished to support this feature, due to hardware constraints. 
%Conclusion
As this was the only reductive change found, we conclude that the evolution of the API was not primarily driven by reductive changes.

\smallskip \noindent
\textbf{Adaptive  }
%Introduction and definition
Adaptive changes are defined to be changes which involve changes to technology or resources used. 
%Data and examples
A single change was identified as being adaptive; the inclusion of new 3\textit{rd} part libraries. Glib existed in version 1.4 of the API its use was increased in version 2.0. Furthermore, other libraries were added for audio and data communication.

This type of change made up 7.7\% of the total changes, however it had a broader impact than other individual changes.

%Conclusion

\smallskip \noindent
\textbf{Performance  }
No identified changes could be related to performance...

\smallskip \noindent
\textbf{Preventive  }
Preventive changes were...

\smallskip \noindent
\textbf{Groomative  }
A couple of groomative changes were found...

\smallskip \noindent
\textbf{Updative  }
Updative changes only affected documentation and not code...

\smallskip \noindent
\textbf{Reformative  }
Blabla...

\smallskip \noindent
\textbf{Evaluative  }
Blabla...

\smallskip \noindent
\textbf{Consultive  }
Blabla...

\smallskip \noindent
\textbf{Training  }
Blabla...

@@

\newpage
\subsection{RQ2 Results}

>>>>

\subsection{API Design} \label{api_design}
API design is notoriously difficult, as a myriad of design and performance decisions must be taken into consideration when creating APIs \cite{afonso2012evaluating} \cite{bloch2008effective} \cite{stylos2006comparing}. Examples of such design decisions include how to structure an object's constructor parameters or if the API should display errors at compilation or at runtime \cite{stylos2006comparing}. More trivial design problems, such as assigning names to API features or correctly naming user-defined types, can have a significant impact on the usability of an API \cite{shi2011empirical}. When facing such design challenges, the following four factors are important to consider: a) The API must be understandable through good documentation, b) the API must not be overly abstract, c) the API must be reusable and d) the API must be easy to learn \cite{shi2011empirical}. One of the most important qualities in an API is that the intent of the API must be clear to the user \cite{stylos2006comparing} \cite{shi2011empirical}. The design decisions reached during development of an API will affect the overall usability of the API. Measuring such an effect can be done by investigating the twelve cognitive dimensions that are impacted by interactions between the API and its users \cite{clarke2004measuring}.

@@

%%Lehman's laws (!!!improve later)
Software evolution theory suggests that changes to software must obey one of eight laws \cite{lehman1980programs}. The laws, known as Lehman's laws, are: 1) Continuing Change, 2) Increasing Complexity, 3) Self Regulation, 4) Conservation of Organisational Stability, 5) Conservation of Familiarity, 6) Continuing Growth, 7) Declining Quality and 8) Feedback System \cite{lehman1980programs}. The laws are said to only apply to E-type systems, which are programs that are ``change prone" and that ``mechanise a human or societal activity" \cite{lehman1980programs}.

@@

\subsection{API Evolution} \label{related_work}
The analysis of APIs in the context of software evolution has primarily focused on APIs that are part of large programming languages, such as Java \cite{hou2011exploring} \cite{shi2011empirical} and Smalltalk \cite{robbes2012developers}. Attempts to uncover the intents behind the changes made in the AWT and Swing Java libraries have also been undertaken \cite{hou2011exploring}, leading to suggestions that the use of a strong architecture is vital for ensuring the successful evolution of an API \cite{hou2011exploring}.

@@

%%Possibly add a paragraph here to

@@

%The case company was selected using convenience sampling \cite{flyvbjerg2006five} and a maximum variation strategy was used \cite{benbasat1987case}. As such, the goal was to include as many companies as possible in this study.

@@

%Reasons for choosing case study
The reason the selected methodology was used is that it is essential to study the phenomenon of API change in its natural context. The applicability of case studies in such scenarios is supported by existing literature \cite{benbasat1987case} \cite{runeson2009guidelines} \cite{yin2013case} \cite{robson2002real}. An alternative approach that was considered was design research, but if a prototype API was to be used instead of one which is tried and tested in an industry setting, the study would lack real-life context \cite{runeson2009guidelines}. An additional motivation for why the case study approach was used is that there is little existing research conducted in the area of motives behind API change and that input from the industry is vital to the success of the research, especially to be able to answer \textbf{RQ2}.

@@

The process used for examining the code and documentation was inspired by grounded theory analysis \cite{seaman1999qualitative}, as it is recommended for hypothesis generating studies \cite{runeson2009guidelines} \cite{seaman1999qualitative}.

@@

\smallskip \noindent
\textbf{Construct validity  } Threats to construct validity have largely been mitigated by structuring the results of the study around already established and accepted software evolution theories \cite{chapin2001types} \cite{lehman1980programs}. Due to restricted access to the company's source code, we were not able to include versions prior to version 1.4 in our analysis. This means that the evolution of the API was studied during a rather short period of its lifetime. It was also not possible to analyse the implementation of the API, something which might have contributed to more accurate predictions of the motives behind the changes. We consider these factors to be the remaining threats to construct validity.

@@

\smallskip \noindent
\textbf{Reliability  } Potential threats to the reliability of our study have, to a great extent, been reduced by conducting the study according to an accredited guide to case studies \cite{runeson2009guidelines}. In addition to that, strategies from additional well-established papers \cite{andersson2007spiral} \cite{seaman1999qualitative} \cite{robson2002real} have been used to increase the trustworthiness of the data gathering and data analysis. The analysis was based around well-defined theories that are cemented in academia, to further increase the reproducibility of the study. The semi-structured nature of the interview, might threaten the reproducibility; however, large parts of it were structured around existing theories.

@@

The majority of changes between version 1.4 and version 2.0 have been additions to functionality of the API. Such additions include added functionality to mechanically and digitally control the camera, to utilise additional storage devices, to allow serial connectivity and to allow audio analysis on the cameras. Pure additions to the API did not have any secondary types of change attributed to them.

Many of the changes found many deprecations which were later re-implemented as new modules. We have grouped these kinds of changes into a single type instead of two. Such examples include the deprecation and re-implementation of an event system, dynamic web page generation and configuration utilities. These kinds of changes often had a secondary type of change that accompanied the primary type of change.

The remainder of the changes were defined as general trends. These changes were not individual, but were instead larger changes that effected the whole API. Such examples include a shift towards using interfaces to access features, improved maintainability, a different error-handling approach and the inclusion of additional 3\textit{rd} party libraries. With the exception of 3\textit{rd} party libraries, the general trends have secondary changes which help to categories what effect the changes had on the API.

@@

\smallskip \noindent
\textbf{Corrective  }
%Define what kinds of changes these are
Corrective changes are identified as types of change that fix broken functionality.
%Data and explanation
No corrective changes were identified between V1.0 and V2.0. There are two main reasons that we did not find any changes of this type. Firstly, the majority of changes were additions or re-implementations of functionality. This meant that most of the existing code was either removed, replaced or deprecated resulting in little to no obvious fixes. Secondly, we only had access to the public facing API and not the implementation itself. We therefore could not inspect the inner workings of the functionality that has remained between versions to determine if any corrective change took place.
%Conclusion
We therefore conclude that the evolution of the API was not driven by corrective changes.
\smallskip \noindent
\textbf{Reductive  }
%Definition
Reductive changes are defined as changes that remove or reduce functionality. This is distinctive from a deprecation as reductive types of change remove functionality completely.
%Data & Explanation
Only a single reductive type of change was identified. This reductive change removed functionality that provided the ability to buffer individual images for closer analysis by a user. Given that it is more common to refactor APIs and not remove functionality \cite{dig2005role, xing2006refactoring} this was a surprising result. 
%Don't know if I should discuss the finding in this much depth here?
%If OK - Talk more about the business side of things, and how we couldn't identify this change ourselves
It was revealed during the interview that the functionality was removed because the case company no longer wished to support this feature, due to hardware constraints. 
%Conclusion
As this was the only reductive change found, we conclude that the evolution of the API was not primarily driven by reductive changes.

\smallskip \noindent
\textbf{Adaptive  }
%Introduction and definition
Adaptive changes are defined to be changes which involve changes to technology or resources used. 
%Data and examples
A single change was identified as being adaptive; the inclusion of new 3\textit{rd} part libraries. Glib existed in version 1.4 of the API its use was increased in version 2.0. Furthermore, other libraries were added for audio and data communication. This type of change made up 7.7\% of the total changes, however it had a broader impact than other individual changes.
%Conclusion 
The evolution of the API has not been primarily driven by adaptive changes. However, more information, possibly in the form of another study, would be required to fully conclude if these types of changes have a greater impact upon the evolution of an API.

\smallskip \noindent
\textbf{Performance  }
%Define
Performance changes are defined to be types of change that intentionally alter system performance.
%Data and examples
No performance types of change were identified in the API. Only having access to the public API and not the implementation meant that we were unable search for performance changes in existing code between the versions. There was no indication from the documentation nor the interview that any change had been performed with the intent of increasing system performance. 
%Conclusion
As such, we conclude that the evolution of the API was not driven by performance changes. A proper benchmark could be performed on sample programs written with both versions of the API to test overall performance. However, this is outside the scope of our study.

\smallskip \noindent
\textbf{Preventive  }
%Define
Preventive changes are defined to be changes that attempt to avoid future maintenance. This should not be confused with changes that attempt to make the current state of the system more maintainable.
%Data and examples
15.8\% of the changes were given the type of Preventive. A new error handling system was one of these changes. It is unclear from version 1.4 what error handling was in place, however, version 2.0 saw the addition of specific error handling modules and functions. According to the interview, the error handling change coincided with the inclusion of more Glib features, which the new error handling system uses. The redesign of the system to use more interfaces was also deemed as a preventive change. Using interfaces to access modules helps to make them more maintainable and more easily extendable, which is a good example of a preventive change.
%Concusion
Preventive types of change appear to drive the evolution of an API more strongly than other types of changes do. This could be due to the fact that as the API grows, the need for higher maintenance also increase, a fact that also seems to be supported by Lehman's laws.

\smallskip \noindent
\textbf{Groomative  }
%Define
Groomative changes are defined to be changes that aim to make the code more maintainable immediately. This should not be confused with preventive changes.
%Data and examples
While none of the change identified were primarily groomative changes, three of them were secondarily categorised as being groomative. This seems to indicate that the groomative nature of a change is often a bonus of other kinds of changes. The move towards using interfaces can be seen as an example of this. While the primary motive of change was to make the API more maintainable in the future, making the API more maintainable immediately was also a reason for the change.
%Conclusion
While not a primary factor for any of the changes, groomative types of change should still be viewed as important for API evolution.

\smallskip \noindent
\textbf{Updative  }
%Define
Updative changes are defined to be changes to documentation and are not changes that affect functionality in any way.

\smallskip \noindent
\textbf{Reformative  }
%Define
Reformative changes are defined to be changes that update the documentation to the stakeholders needs.

\smallskip \noindent
\textbf{Evaluative, Consultive and Training  }
%Define
Blabla...

@@

\subsection{RQ2 Results}
